---
title: "Report"
output:
  html_document:
    df_print: paged
  pdf_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(SelectionSimulator)
library(ggplot2); theme_set(theme_minimal())
source("./simulations/normal_test.R")
```

# Introduction
Quote paper

David Hamilton (1987) Sometimes R 2 > r 2 yx 1 + r 2 yx 2 : Correlated Variables are Not Always Redundant, The American Statistician, 41:2, 129-132, DOI: 10.1080/00031305.1987.10475460

Explain problem wrt stepwise variable selection


# Algorithm

Explain algorithm and hyperparameters, including time and space complexity


# Simulations
How it was simulated/methods, then results

## Two negatively correlated
Covariance matrix = 
$\begin{bmatrix} 
1   & 0   & 0 & 0 & 0 \\
0   & 1   & 0 & 0 & 0 \\
0   & 0   & 1 & 0 & 0 \\
0   & 0   & 0 & 1 & -0.8 \\
0   & 0   & 0 & -0.8 & 1
\end{bmatrix}$

Model coefficients = $\begin{bmatrix} 0 \\ 0 \\ 0 \\ 1 \\ 1 \end{bmatrix}$

n = 100
```{r}
cov_mat_twoNeg <- rbind( c( 1.00,  0.00,  0.00, 0.00, 0.00),
                    c( 0.00,  1.00,  0.00, 0.00, 0.00),
                    c( 0.00,  0.00,  1.00, 0.00, 0.00),
                    c( 0.00,  0.00,  0.00, 1.00, -0.8),
                    c( 0.00,  0.00,  0.00, -0.8, 1.00))

coefficients_twoNeg <- c(V1 = 0, V2 = 0, V3 = 0, V4 = 1, V5 = 1)

test_twoNeg <- normal_test(cov_mat_twoNeg, coefficients_twoNeg, intercept = 9, 
                        num_observations = 100, num_simulations = 500,
                        file_name = "twoNeg.RDS")
```


```{r}
betas_heat_map(test_twoNeg)
#inclusion_order(test_twoNeg)
test_mse_plotter(test_twoNeg)
training_mse_plotter(test_twoNeg)
```

```{r, ascii = TRUE}
test_mse_tables(test_twoNeg)
training_mse_tables(test_twoNeg)
```

### Three Correlated, Step3

Covariance matrix = 
$\begin{bmatrix} 
1     & -0.8  & 0.25  & 0 & 0 \\
-0.8  & 1     & -0.75 & 0 & 0 \\
0.25  & -0.75 & 1     & 0 & 0 \\
0     & 0     & 0     & 1 & 0 \\
0     & 0     & 0     & 0 & 1
\end{bmatrix}$

Model coefficients = 
$\begin{bmatrix} \text{V1} =1 \\ \text{V2} =1 \\
\text{V3} =1 \\ \text{V4} =0 \\ \text{V5} = 0 \end{bmatrix}$

```{r}
cov_mat_threeNeg_5 <- rbind( c(1.00, -0.80,  0.25, 0.00, 0.00),
                    c(-0.80,  1.00, -0.75, 0.00, 0.00),
                    c( 0.25, -0.75,  1.00, 0.00, 0.00),
                    c( 0.00,  0.00,  0.00, 1.00, 0),
                    c( 0.00,  0.00,  0.00, 0, 1.00))

threeNeg_5_coeff <- c(V1 = 1, V2 = 1, V3 = 1, V4 = 0, V5 =0)

test_threeNeg_10 <- normal_test(cov_mat_threeNeg_5, 
                                coefficients = threeNeg_5_coeff, intercept = 9, 
                        num_observations = 100, num_simulations = 500,
                        file_name = "threeNeg.RDS")
```

```{r}
betas_heat_map(test_threeNeg_10)
inclusion_order(test_threeNeg_10)
test_mse_plotter(test_threeNeg_10)
training_mse_plotter(test_threeNeg_10)
test_mse_tables(test_threeNeg_10)
training_mse_tables(test_threeNeg_10)
```


## Three correlated

Covariance matrix = 
$\begin{bmatrix} 
1   & -0.8& 0.25& 0   & 0   & 0 & 0 & 0 & 0 & 0\\
-0.8& 1   &-0.75& 0   & 0   & 0 & 0 & 0 & 0 & 0\\
0.25&-0.75& 1   & 0   & 0   & 0 & 0 & 0 & 0 & 0\\
0   & 0   & 0   & 1   &-0.75& 0 & 0 & 0 & 0 & 0\\
0   & 0   & 0   &-0.75& 1   & 0 & 0 & 0 & 0 & 0 \\
0   & 0   & 0   & 0   & 0   & 1 & -0.8 & 0.25 & 0 & 0 \\
0   & 0   & 0   & 0   & 0   &-0.8& 1 & -0.75 & 0 & 0 \\
0   & 0   & 0   & 0   & 0   &0.25& -0.75 & 1 & 0 & 0 \\
0   & 0   & 0   & 0   & 0   & 0 & 0 & 0 & 1 & -0.75 \\
0   & 0   & 0   & 0   & 0   & 0 & 0 & 0 & -0.75 & 1
\end{bmatrix}$

Coefficients = 
```{r eval=FALSE, include=FALSE}

test_threeNeg_10 <- normal_test(cov_mat_twoNeg, coefficients_twoNeg, intercept = 9, 
                        num_observations = 100, num_simulations = 500,
                        file_name = "twoNeg.RDS")

cov_mat_5 <- rbind( c(1.00, -0.80,  0.25, 0.00, 0.00),
                    c(-0.80,  1.00, -0.75, 0.00, 0.00),
                    c( 0.25, -0.75,  1.00, 0.00, 0.00),
                    c( 0.00,  0.00,  0.00, 1.00, -0.75),
                    c( 0.00,  0.00,  0.00, -0.75, 1.00))

coefficients_1 <- c(replicate(5, 0), replicate(5, 1))
cov_mat_10 <- as.matrix(Matrix::bdiag(replicate(2, cov_mat_5, simplify = FALSE)))
predict_gen <- SelectionSimulator::NormalPredictorsGenerator$new(
  num_observations = 100,
  num_predictors = 10,
  norm_rand_var_sd = 1,
  covariance_matrix = cov_mat_10)

names(coefficients_1) <- sapply(seq.int(10), function(x) paste0("V", x))

response_calc_1 <- SelectionSimulator::LinearNormalResponseCalculator$new(
  norm_rand_var_sd = 1,
  coefficients = coefficients_1,
  intercept = 9)

sim_data_gen <- SelectionSimulator::SimulationDataGenerator$new(
  predict_gen, response_calc_1)


constructors <- c(
  Step = SelectionSimulator::StepLmWrapper$new,
  Step2 = SelectionSimulator::Step2$new,
  Step3 = SelectionSimulator::Step3$new)

test_obj <- SelectionSimulator::Simulation$new(sim_data_gen, constructors, 500, 1)

test_obj$simulate()
```


```{r eval=FALSE, include=FALSE}
betas_heat_map(test_obj)
test_mse_plotter(test_obj)
training_mse_plotter(test_obj)
test_mse_tables(test_obj)
training_mse_tables(test_obj)
#inclusion_order(test_obj)
```

## Three correlated

Covariance matrix = 
$\begin{bmatrix} 
1 & -0.8 & 0.25 & 0 & 0 \\
-0.8 & 1 & -0.75 & 0 & 0 \\
0.25 & -0.75 & 1 & 0 & 0 \\
0 & 0 & 0 & 1 & -0.75 \\
0 & 0 & 0 & -0.75 & 1
\end{bmatrix}$

Coefficients = $\begin{bmatrix} 0 \\ 1 \\ 0 \\ 1 \\ 0 \end{bmatrix}$
```{r eval=FALSE, include=FALSE}
cov_mat_5 <- rbind( c(1.00, -0.80,  0.25, 0.00, 0.00),
                    c(-0.80,  1.00, -0.75, 0.00, 0.00),
                    c( 0.25, -0.75,  1.00, 0.00, 0.00),
                    c( 0.00,  0.00,  0.00, 1.00, -0.75),
                    c( 0.00,  0.00,  0.00, -0.75, 1.00))

coefficients_1 <- c(V1 = 0, V2 = 1, V3 = 0, V4 = 1, V5 = 0)

predict_gen <- NormalPredictorsGenerator$new(num_observations = 50,
                                             num_predictors = 5,
                                             norm_rand_var_sd = 1,
                                             covariance_matrix = cov_mat_5)

response_calc_1 <- LinearNormalResponseCalculator$new(norm_rand_var_sd = 1,
                                                      coefficients = coefficients_1,
                                                      intercept = 9)

sim_data_gen <- SimulationDataGenerator$new(predict_gen, response_calc_1)


constructors <- c(Step = StepLmWrapper$new, Step2 = Step2$new, Step3 = Step3$new)

test_obj <- SelectionSimulator::Simulation$new(sim_data_gen, constructors, 500, 1)

test_obj$simulate()

```

```{r eval=FALSE, include=FALSE}
betas_heat_map(test_obj)
#inclusion_order(test_obj)
test_mse_plotter(test_obj)
training_mse_plotter(test_obj)
test_mse_tables(test_obj)
training_mse_tables(test_obj)
```
